### 2014-04-18 article.tex
Rename report.tex to article.tex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% American Geophysical Union (AGU)
% LaTeX Template
% Version 1.0 (3/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% The AGUTeX class and agu-ps referencing style were created and are owned 
% by AGU: http://publications.agu.org/author-resource-center/author-guide/latex-formatting-toolkit/
%
% This template has been modified from the blank AGU template to include
% examples of how to insert content and drastically change commenting. The
% structural integrity is maintained as in the original blank template.
%
% Important notes: 
% This template retains extensive commenting from the AGU template. It is heavily 
% advised you read these comments and follow them in order to insure a speedy 
% submission process.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified March 2013
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Step 1: Set the \documentclass

% There are two options for article format: two column (default) and draft.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.

% Choose the journal abbreviation for the journal you are submitting to:

% jgrga	JOURNAL OF GEOPHYSICAL RESEARCH
% gbc	GLOBAL BIOCHEMICAL CYCLES
% grl		GEOPHYSICAL RESEARCH LETTERS
% pal	PALEOCEANOGRAPHY
% ras	RADIO SCIENCE
% rog	REVIEWS OF GEOPHYSICS
% tec	TECTONICS
% wrr	WATER RESOURCES RESEARCH
% gc		GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw	SPACE WEATHER
% ms	JAMES
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

%\documentclass[grl]{AGUTeX}
\documentclass[draft,grl]{AGUTeX}

% To create numbered lines:

% If you don't already have lineno.sty, you can download it from http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/ (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN). Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno} and \linenumbers*[1]command, below:

\usepackage{mhchem}
\usepackage{lineno}
\linenumbers*[1]

%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables

% DO NOT USE \psfrag or \subfigure commands.

%  Figures and tables should be placed AT THE END OF THE ARTICLE, after the references.

%  Uncomment the following command to include .eps files (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
\usepackage{graphicx}

% Substitute one of the following for [dvips] above if you are using a different driver program and want to proof your illustrations on your machine:
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]

%  Uncomment the following command to allow illustrations to print when using Draft:
\setkeys{Gin}{draft=false}

% See how to enter figures and tables at the end of the article, after references.

%----------------------------------------------------------------------------------------
%	RUNNING HEAD AND CORRESPONDING AUTHOR
%----------------------------------------------------------------------------------------

% Author names in capital letters:
\authorrunninghead{BTTM8}
%------------------------------------------------

% Shorter version of title entered in capital letters:
\titlerunninghead{CLIMATE SENSITIVITY USING HADAM3}

%------------------------------------------------

% Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: XXX, MSc Student, Department of Geography, UCL (XXX@ucl.ac.uk)}

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\title{Analysis of Climate Sensitivity using HadAM3}

%----------------------------------------------------------------------------------------
%	AUTHORS AND AFFILIATIONS
%----------------------------------------------------------------------------------------

% Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote; matching \altaffiltext will appear at bottom of page.

\authors{Candidate Number: BTTM8\altaffilmark{1}}

\altaffiltext{1}{Department of Geography, UCL}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% Do NOT include any \begin...\end commands within the body of the abstract.

\begin{abstract}

\ce{CO2} ipsum dolor sit amet, consectetur adipiscing elit. Sed vehicula metus sapien. Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu. Vestibulum ultrices eros sed odio tincidunt bibendum. Pellentesque fermentum ante vel nulla commodo fermentum. Vestibulum in augue sit amet libero viverra accumsan eu at magna. Sed at ligula quis nibh pharetra facilisis non eu libero. Suspendisse non quam sit amet massa luctus interdum sit amet in purus. Integer id orci elit, vitae sollicitudin lectus.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed vehicula metus sapien. Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu. Vestibulum ultrices eros sed odio tincidunt bibendum. Pellentesque fermentum ante vel nulla commodo fermentum. Vestibulum in augue sit amet libero viverra accumsan eu at magna. Sed at ligula quis nibh pharetra facilisis non eu libero. Suspendisse non quam sit amet massa luctus interdum sit amet in purus. Integer id orci elit, vitae sollicitudin lectus.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENT
%----------------------------------------------------------------------------------------

% The body of the article must start with a \begin{article} command
% \end{article} must follow the references section, before the figures and tables.

\begin{article}

\section{Introduction}

Nam fermentum sapien at enim varius consectetur. Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae. Integer lacinia purus vel metus tempus suscipit. Curabitur ac sapien quis mauris euismod commodo. Sed pharetra sem elit. Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor. Donec convallis ligula vitae sem viverra fermentum. Mauris in ullamcorper erat. Donec ultrices tempus nibh quis vestibulum. This statement requires citation \citep{bro1967}. This one is an in-text citation because the authors of \citet{Englemann1986} are specifically mentioned.

Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus. Aliquam a tellus nec felis sagittis consequat. Quisque convallis imperdiet neque a tempor. Nulla non erat urna. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Suspendisse viverra tortor pretium orci lacinia eleifend. Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem. Vestibulum pharetra malesuada egestas. Sed non orci sit amet erat suscipit fringilla in et diam. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla. Aenean porta rhoncus suscipit.

%------------------------------------------------

\section{Materials and Methods}

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Mauris eu sapien nunc, sit amet accumsan dui. Nulla ac diam ut nunc placerat semper eget et libero. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Cras hendrerit ullamcorper sapien vitae luctus. Quisque vel diam massa. Vestibulum dui nibh, facilisis vel vestibulum eu, viverra in quam.

\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%------------------------------------------------

\section{Results}

Referencing equation \ref{eq:emc}. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Referencing Table \ref{sampletable}. Referencing Figure \ref{placeholder}.

\subsection{Simulations}

\subsubsection{Simulation 1}

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

\subsubsection{Simulation 2}

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

\subsection{Real Data}

Aliquam interdum pellentesque scelerisque. Sed tincidunt suscipit purus, id aliquet nulla vehicula quis. Duis sed nisl lorem. Vivamus erat ante, dignissim et aliquam vel, adipiscing vitae magna. Cras id dapibus metus. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Proin ut lectus ut nisi congue ullamcorper. Ut ac turpis ligula. Sed faucibus bibendum nunc eget gravida.

%------------------------------------------------

\section{Discussion}

Nam fermentum sapien at enim varius consectetur. Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae. Integer lacinia purus vel metus tempus suscipit. Curabitur ac sapien quis mauris euismod commodo. Sed pharetra sem elit. Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor. Donec convallis ligula vitae sem viverra fermentum. Mauris in ullamcorper erat. Donec ultrices tempus nibh quis vestibulum.

Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus. Aliquam a tellus nec felis sagittis consequat. Quisque convallis imperdiet neque a tempor. Nulla non erat urna. Mauris vel lorem magna, tristique auctor ipsum. Aliquam pharetra eleifend massa. Donec porttitor sagittis luctus. Aliquam pretium luctus leo quis congue. Morbi vel felis mi. Suspendisse viverra tortor pretium orci lacinia eleifend. Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem. Vestibulum pharetra malesuada egestas. Sed non orci sit amet erat suscipit fringilla in et diam. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla. Aenean porta rhoncus suscipit.

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

Suspendisse viverra eleifend nulla at facilisis. Nullam eget tellus orci. Cras sit amet lorem velit. Maecenas rhoncus pellentesque orci eget vulputate. Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque. In ac enim elit, sit amet luctus ante. Cras feugiat commodo lectus, vitae convallis dui sagittis id. In in tellus lacus, sed lobortis eros. Phasellus sit amet eleifend velit. Duis ornare dapibus porttitor. Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat. Proin elementum mi vel lectus suscipit fringilla. Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi. Maecenas ut massa aliquet purus eleifend vehicula in a nisi. Fusce molestie cursus lacinia.

%----------------------------------------------------------------------------------------
%	APPENDICES (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here

% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing  \appendix

% \section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title

\appendix

\section{Appendix Title}

Vivamus magna enim, aliquet id cursus a, pharetra ut purus. Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum. Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida. Etiam sagittis commodo neque eget placerat. Sed et nisi faucibus metus interdum adipiscing id nec lacus. Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam. Phasellus cursus semper viverra. Proin magna tortor, blandit in ultricies id, facilisis at nibh. Proin eu neque est. Etiam euismod auctor ante. Mauris mauris sem, tincidunt a placerat rutrum, porta id est. Aenean non velit porta eros condimentum facilisis at in nibh. Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor. Etiam ac leo at ipsum tincidunt consequat ac non sapien. Aenean sed leo diam, venenatis pharetra odio.

%----------------------------------------------------------------------------------------
%	GLOSSARY OR NOTATION (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgments}
This work was partially supported by a grant from the Spanish Ministry of Science and Technology.
\end{acknowledgments}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}

% Or,

% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.

% Follow these steps:
% 1. Run LaTeX on your LaTeX file.

% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.

% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.

% 4. Comment out the old \bibliographystyle and \bibliography commands.

% 5. Run LaTeX on your new file before submitting.

% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

%\bibliographystyle{agufull08}
%\bibliography{sample-bib-08}
\begin{thebibliography}{2}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Brown}(1967)}]{bro1967}
Brown, R. J.~E. (1967), Permafrost in {Canada}, \textit{Map 1246A}, Geol. Surv.
  of Can., Ottawa, Ont.

\bibitem[{\textit{Englemann et~al.}(1986)\textit{Englemann, Perkins, Hagan, and
  Haller}}]{Englemann1986}
Englemann, R.~J., R.~W. Perkins, D.~I. Hagan, and W.~A. Haller (1986), Washout
  coefficients for selected gases and particulates, paper presented at 59th
  Annual Meeting, Air Pollut. Control Agency, San Francisco, Calif., 20--24
  June.

\end{thebibliography}

% Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

\end{article}

%----------------------------------------------------------------------------------------
%	FIGURES AND TABLES
%----------------------------------------------------------------------------------------

%% Enter Figures and Tables here:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information should be placed in footnotes below the table.
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}
% \noindent\includegraphics[width=20pc]{samplefigure.eps}
% \caption{Caption text here}
% \label{figure_label}
% \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% See below for how to make sideways figures or tables.

\newpage

\begin{figure}
\includegraphics[width=0.4\linewidth]{placeholder.jpg}
\caption{Figure caption}\label{placeholder}
\end{figure}

\begin{table}
\caption{Table caption}\label{sampletable}
\begin{tabular}{l l l}
\hline
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\hline
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\hline
\end{tabular}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with the first character following the equal sign on the first line.

% Use an \hspace{} command to insert horizontal space into your equation if necessary. Place an appropriate unit of measure between the curly braces, e.g. \hspace{1in}; you may have to experiment to achieve the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}

### 2014-04-18 article.tex
Add in a propert (read from .bib files) bibliography/ref section.
Add in a handy little tool for gen/view/cleaning up latex files.

@@
 
+% Does not apply to me as I'll just be handing in the PDF.
 % Either type in your references using
@@
 
-%\bibliographystyle{agufull08}
-%\bibliography{sample-bib-08}
-\begin{thebibliography}{2}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
+% BIB_RM
+\bibliographystyle{agufull08}
+\bibliography{sample-bib-08}
+% END_BIB_RM
 
-\bibitem[{\textit{Brown}(1967)}]{bro1967}
-Brown, R. J.~E. (1967), Permafrost in {Canada}, \textit{Map 1246A}, Geol. Surv.
-  of Can., Ottawa, Ont.
+% BIB_POS
 
-\bibitem[{\textit{Englemann et~al.}(1986)\textit{Englemann, Perkins, Hagan, and
-  Haller}}]{Englemann1986}
-Englemann, R.~J., R.~W. Perkins, D.~I. Hagan, and W.~A. Haller (1986), Washout
-  coefficients for selected gases and particulates, paper presented at 59th
-  Annual Meeting, Air Pollut. Control Agency, San Francisco, Calif., 20--24
-  June.
-
-\end{thebibliography}
+%\begin{thebibliography}{2}
+%\providecommand{\natexlab}[1]{#1}
+%\expandafter\ifx\csname urlstyle\endcsname\relax
+%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+%  \urlstyle{rm}\Url}\fi
+%
+%\bibitem[{\textit{Brown}(1967)}]{bro1967}
+%Brown, R. J.~E. (1967), Permafrost in {Canada}, \textit{Map 1246A}, Geol. Surv.
+%  of Can., Ottawa, Ont.
+%
+%\bibitem[{\textit{Englemann et~al.}(1986)\textit{Englemann, Perkins, Hagan, and
+%  Haller}}]{Englemann1986}
+%Englemann, R.~J., R.~W. Perkins, D.~I. Hagan, and W.~A. Haller (1986), Washout
+%  coefficients for selected gases and particulates, paper presented at 59th
+%  Annual Meeting, Air Pollut. Control Agency, San Francisco, Calif., 20--24
+%  June.
+%
+%\end{thebibliography}
 

### 2014-04-18 gen_pdf.py

@@ -1,0 +1,66 @@
+#!/usr/bin/python
+import argparse
+from glob import glob
+import os, sys
+from subprocess import call, Popen
+
+OUTPUT_DIR = 'output'
+
+def clean(output_dir):
+    exts = ['.dvi', '.aux', '.blg', '.pdf', '.bcf', '.log', '.xml', '.bbl']
+    for ext in exts:
+        search_string = '%s/*%s'%(output_dir, ext)
+        filenames = glob(search_string)
+        for filename in filenames:
+            os.remove(filename)
+
+def make_pdf(output_dir, base_filename):
+    pdf_cmd = ('pdflatex --output-directory=%s %s.tex'%(output_dir, base_filename))
+    bibtex_cmd = ('bibtex %s/%s.aux'%(output_dir, base_filename))
+    call(pdf_cmd.split(' '))
+    call(bibtex_cmd.split(' '))
+
+def view(output_dir, base_filename):
+    pdf_view_cmd = 'evince %s/%s.pdf'%(output_dir, base_filename)
+    Popen(pdf_view_cmd.split(' '))
+
+def main(args):
+    output_dir = args.output_dir if args.output_dir else OUTPUT_DIR
+    if args.clean:
+        clean(output_dir)
+    if args.make:
+        if not os.path.exists(output_dir):
+            os.mkdir(output_dir)
+        make_pdf(output_dir, args.base_filename)
+    if args.view:
+        view(output_dir, args.base_filename)
+    return 1
+
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser()
+    parser.add_argument('-o', '--output-dir')
+    parser.add_argument('-m', '--make', action='store_true')
+    parser.add_argument('-c', '--clean', action='store_true')
+    parser.add_argument('-v', '--view', action='store_true')
+    parser.add_argument('input_filename', nargs='?')
+
+    args = parser.parse_args()
+    if args.clean:
+        main(args)
+    elif args.make:
+        if not args.input_filename or os.path.splitext(args.input_filename)[-1] != '.tex':
+            print('Please enter a tex filename')
+        else:
+            if not os.path.exists(args.input_filename):
+                print('File %s does not exist'%args.input_filename)
+            else:
+                args.base_filename = '.'.join(os.path.splitext(args.input_filename)[:-1])
+                main(args)
+    elif args.view:
+        if not args.input_filename or os.path.splitext(args.input_filename)[-1] != '.pdf':
[…]

### 2014-04-25 article.tex
Change from Geophys. research letter tpl to custom one (based on geogg101)

--- 2014-04-18 article.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% American Geophysical Union (AGU)
-% LaTeX Template
-% Version 1.0 (3/6/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% The AGUTeX class and agu-ps referencing style were created and are owned 
-% by AGU: http://publications.agu.org/author-resource-center/author-guide/latex-formatting-toolkit/
-%
-% This template has been modified from the blank AGU template to include
-% examples of how to insert content and drastically change commenting. The
-% structural integrity is maintained as in the original blank template.
-%
-% Important notes: 
-% This template retains extensive commenting from the AGU template. It is heavily 
-% advised you read these comments and follow them in order to insure a speedy 
-% submission process.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{article}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
-% Modified March 2013
-%
-% This template includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements.
-%
-% PLEASE DO NOT USE YOUR OWN MACROS
-% DO NOT USE \newcommand, \renewcommand, or \def.
-%
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% All questions should be e-mailed to latex@agu.org.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Some useful packages.
+\usepackage{amsmath}
+\usepackage{siunitx}
+\usepackage{graphicx}
+\usepackage{verbatim}
+\usepackage{mhchem}
 
-% Step 1: Set the \documentclass
+% Reduces margins substantially.
+\usepackage{geometry}
+\newgeometry{margin=2.5cm}
 
-% There are two options for article format: two column (default) and draft.
+% Allows headers and footers.
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+% Get rid of annoying line under header.
+\renewcommand{\headrulewidth}{0pt}
 
-% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-% The draft option produces double spaced output.
+\lhead{}
+\chead{BTTM8 - CLIMATE MODELLING - USING HADAM3 TO MODEL CLIMATE CHANGE}
+\rhead{}
+%\lfoot{}
+%\cfoot{}
+%\rfoot{}
 
-% Choose the journal abbreviation for the journal you are submitting to:
-
-% jgrga	JOURNAL OF GEOPHYSICAL RESEARCH
-% gbc	GLOBAL BIOCHEMICAL CYCLES
-% grl		GEOPHYSICAL RESEARCH LETTERS
-% pal	PALEOCEANOGRAPHY
-% ras	RADIO SCIENCE
-% rog	REVIEWS OF GEOPHYSICS
-% tec	TECTONICS
-% wrr	WATER RESOURCES RESEARCH
-% gc		GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
-% sw	SPACE WEATHER
-% ms	JAMES
-%
-%
-%
-% (If you are submitting to a journal other than jgrga,
-% substitute the initials of the journal for "jgrga" below.)
-
-%\documentclass[grl]{AGUTeX}
-\documentclass[draft,grl]{AGUTeX}
-
-% To create numbered lines:
-
-% If you don't already have lineno.sty, you can download it from http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/ (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
-Take care that you always use the latest version.
-
-% To activate the commands, uncomment \usepackage{lineno} and \linenumbers*[1]command, below:
-
-\usepackage{mhchem}
-\usepackage{lineno}
-\linenumbers*[1]
-
-%  To add line numbers to lines with equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figures and Tables
-
-% DO NOT USE \psfrag or \subfigure commands.
-
-%  Figures and tables should be placed AT THE END OF THE ARTICLE, after the references.
-
-%  Uncomment the following command to include .eps files (comment out this line for draft format):
-%\usepackage[dvips]{graphicx}
-\usepackage{graphicx}
-
-% Substitute one of the following for [dvips] above if you are using a different driver program and want to proof your illustrations on your machine:
-% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
-% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
-% [oztex], [textures]
-
-%  Uncomment the following command to allow illustrations to print when using Draft:
-\setkeys{Gin}{draft=false}
-
-% See how to enter figures and tables at the end of the article, after references.
-
-%----------------------------------------------------------------------------------------
-%	RUNNING HEAD AND CORRESPONDING AUTHOR
-%----------------------------------------------------------------------------------------
-
-% Author names in capital letters:
-\authorrunninghead{BTTM8}
-%------------------------------------------------
-
-% Shorter version of title entered in capital letters:
-\titlerunninghead{CLIMATE SENSITIVITY USING HADAM3}
-
-%------------------------------------------------
-
-% Corresponding author mailing address and e-mail address:
-\authoraddr{Corresponding author: XXX, MSc Student, Department of Geography, UCL (XXX@ucl.ac.uk)}
-
-%----------------------------------------------------------------------------------------
+\usepackage[backend=bibtex,style=authoryear,sorting=nyt,dashed=false]{biblatex}
+\bibliography{references/references}
+\renewcommand*{\nameyeardelim}{\addcomma\space}
 
@@
 
-%----------------------------------------------------------------------------------------
-%	TITLE
-%----------------------------------------------------------------------------------------
+\part*{Using HadAM3 to model the effects of climate change}
 
-\title{Analysis of Climate Sensitivity using HadAM3}
+\section*{Abstract}
 
-%----------------------------------------------------------------------------------------
-%	AUTHORS AND AFFILIATIONS
-%----------------------------------------------------------------------------------------
-
-% Use \author{\altaffilmark{}} and \altaffiltext{}
-
-% \altaffilmark will produce footnote; matching \altaffiltext will appear at bottom of page.
-
-\authors{Candidate Number: BTTM8\altaffilmark{1}}
-
-\altaffiltext{1}{Department of Geography, UCL}
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-% Do NOT include any \begin...\end commands within the body of the abstract.
-
-\begin{abstract}
-
-\ce{CO2} ipsum dolor sit amet, consectetur adipiscing elit.
-Sed vehicula metus sapien.
-Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu.
-Vestibulum ultrices eros sed odio tincidunt bibendum.
-Pellentesque fermentum ante vel nulla commodo fermentum.
-Vestibulum in augue sit amet libero viverra accumsan eu at magna.
-Sed at ligula quis nibh pharetra facilisis non eu libero.
-Suspendisse non quam sit amet massa luctus interdum sit amet in purus.
-Integer id orci elit, vitae sollicitudin lectus.
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed vehicula metus sapien.
-Suspendisse pulvinar, felis ut hendrerit aliquet, dui nisi bibendum erat, fermentum mattis enim nibh id arcu.
-Vestibulum ultrices eros sed odio tincidunt bibendum.
-Pellentesque fermentum ante vel nulla commodo fermentum.
-Vestibulum in augue sit amet libero viverra accumsan eu at magna.
-Sed at ligula quis nibh pharetra facilisis non eu libero.
-Suspendisse non quam sit amet massa luctus interdum sit amet in purus.
-Integer id orci elit, vitae sollicitudin lectus.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENT
-%----------------------------------------------------------------------------------------
-
-% The body of the article must start with a \begin{article} command
-% \end{article} must follow the references section, before the figures and tables.
-
-\begin{article}
+\ce{CO2} climate sensitivity (\SI{}{W.m^{-2}.K^{-1}})
 
@@
 
-Nam fermentum sapien at enim varius consectetur.
-Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae.
-Integer lacinia purus vel metus tempus suscipit.
-Curabitur ac sapien quis mauris euismod commodo.
-Sed pharetra sem elit.
-Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor.
-Donec convallis ligula vitae sem viverra fermentum.
-Mauris in ullamcorper erat.
-Donec ultrices tempus nibh quis vestibulum.
-This statement requires citation \citep{bro1967}.
-This one is an in-text citation because the authors of \citet{Englemann1986} are specifically mentioned.
+According to \textcite{gregory2004new}, you can never \dots
 
-Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus.
-Aliquam a tellus nec felis sagittis consequat.
-Quisque convallis imperdiet neque a tempor.
-Nulla non erat urna.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Suspendisse viverra tortor pretium orci lacinia eleifend.
-Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem.
-Vestibulum pharetra malesuada egestas.
-Sed non orci sit amet erat suscipit fringilla in et diam.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla.
-Aenean porta rhoncus suscipit.
+\section{Methods}
 
-%------------------------------------------------
-
-\section{Materials and Methods}
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
-
-\begin{equation}
-\label{eq:emc}
-e = mc^2
-\end{equation}
-
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Mauris eu sapien nunc, sit amet accumsan dui.
-Nulla ac diam ut nunc placerat semper eget et libero.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Cras hendrerit ullamcorper sapien vitae luctus.
-Quisque vel diam massa.
-Vestibulum dui nibh, facilisis vel vestibulum eu, viverra in quam.
-
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%------------------------------------------------
+\subsection{Model}
 
@@
 
-Referencing equation \ref{eq:emc}.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Referencing Table \ref{sampletable}.
-Referencing Figure \ref{placeholder}.
-
-\subsection{Simulations}
-
-\subsubsection{Simulation 1}
-
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-\subsubsection{Simulation 2}
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
-
-\subsection{Real Data}
-
-Aliquam interdum pellentesque scelerisque.
-Sed tincidunt suscipit purus, id aliquet nulla vehicula quis.
-Duis sed nisl lorem.
-Vivamus erat ante, dignissim et aliquam vel, adipiscing vitae magna.
-Cras id dapibus metus.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Proin ut lectus ut nisi congue ullamcorper.
-Ut ac turpis ligula.
-Sed faucibus bibendum nunc eget gravida.
-
-%------------------------------------------------
+\begin{figure}[hbp]
+    \centering
+    \includegraphics[width=\textwidth]{figures/figure_1}
+    \caption{Caption.}
+    \label{fig:figure_1}
+\end{figure}
 
@@
 
-Nam fermentum sapien at enim varius consectetur.
-Quisque lobortis imperdiet mauris, et accumsan libero vulputate vitae.
-Integer lacinia purus vel metus tempus suscipit.
-Curabitur ac sapien quis mauris euismod commodo.
-Sed pharetra sem elit.
-Fusce ultrices, mauris eu fermentum tempor, tellus sem ornare lectus, in convallis nunc urna id dolor.
-Donec convallis ligula vitae sem viverra fermentum.
-Mauris in ullamcorper erat.
-Donec ultrices tempus nibh quis vestibulum.
-
-Praesent volutpat, nibh in dignissim commodo, tellus justo consequat erat, vel consequat mi arcu vel lectus.
-Aliquam a tellus nec felis sagittis consequat.
-Quisque convallis imperdiet neque a tempor.
-Nulla non erat urna.
-Mauris vel lorem magna, tristique auctor ipsum.
-Aliquam pharetra eleifend massa.
-Donec porttitor sagittis luctus.
-Aliquam pretium luctus leo quis congue.
-Morbi vel felis mi.
-Suspendisse viverra tortor pretium orci lacinia eleifend.
-Phasellus aliquam, nunc eu cursus feugiat, erat odio porttitor libero, quis accumsan orci ipsum ut lorem.
-Vestibulum pharetra malesuada egestas.
-Sed non orci sit amet erat suscipit fringilla in et diam.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc ut rhoncus nulla.
-Aenean porta rhoncus suscipit.
-
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-Suspendisse viverra eleifend nulla at facilisis.
-Nullam eget tellus orci.
-Cras sit amet lorem velit.
-Maecenas rhoncus pellentesque orci eget vulputate.
-Phasellus massa nisi, mattis nec elementum accumsan, blandit non neque.
-In ac enim elit, sit amet luctus ante.
-Cras feugiat commodo lectus, vitae convallis dui sagittis id.
-In in tellus lacus, sed lobortis eros.
-Phasellus sit amet eleifend velit.
-Duis ornare dapibus porttitor.
-Maecenas eros velit, dignissim at egestas in, tincidunt lacinia erat.
-Proin elementum mi vel lectus suscipit fringilla.
-Mauris justo est, ullamcorper in rutrum interdum, accumsan eget mi.
-Maecenas ut massa aliquet purus eleifend vehicula in a nisi.
-Fusce molestie cursus lacinia.
-
-%----------------------------------------------------------------------------------------
-%	APPENDICES (OPTIONAL)
-%----------------------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-
-% \appendix resets counters and redefines section heads
-% but doesn't print anything.
-% After typing  \appendix
-
-% \section{Here Is Appendix Title}
-% will show
-% Appendix A: Here Is Appendix Title
+\section{Conclusions}
 
@@
 
-\section{Appendix Title}
+\section{Required output}
 
-Vivamus magna enim, aliquet id cursus a, pharetra ut purus.
-Phasellus suscipit nisi iaculis mi vulputate id interdum velit dictum.
-Nam ullamcorper elit in lectus ultrices vitae volutpat massa gravida.
-Etiam sagittis commodo neque eget placerat.
-Sed et nisi faucibus metus interdum adipiscing id nec lacus.
-Donec ipsum diam, malesuada at euismod consectetur, placerat quis diam.
-Phasellus cursus semper viverra.
-Proin magna tortor, blandit in ultricies id, facilisis at nibh.
-Proin eu neque est.
-Etiam euismod auctor ante.
-Mauris mauris sem, tincidunt a placerat rutrum, porta id est.
-Aenean non velit porta eros condimentum facilisis at in nibh.
-Etiam cursus purus ut orci rhoncus sit amet semper eros porttitor.
-Etiam ac leo at ipsum tincidunt consequat ac non sapien.
-Aenean sed leo diam, venenatis pharetra odio.
-
-%----------------------------------------------------------------------------------------
-%	GLOSSARY OR NOTATION (OPTIONAL)
-%----------------------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary or Notation section, goes here
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-% \section*{Glossary}
-% \paragraph{Term}
-% Term Definition here
-%
-%%%%%%%%%%%%%%
-% Notation -- End each entry with a period.
-% \begin{notation}
-% Term & definition.\\
-% Second term & second definition.\\
-% \end{notation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
-\begin{acknowledgments}
-This work was partially supported by a grant from the Spanish Ministry of Science and Technology.
-\end{acknowledgments}
-
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-
-% Or,
-
-% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
-% file and copy the contents into your paper here.
-
-% Follow these steps:
-% 1. Run LaTeX on your LaTeX file.
-
-% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
-% file.
-
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-
-% 4. Comment out the old \bibliographystyle and \bibliography commands.
-
-% 5. Run LaTeX on your new file before submitting.
-
-% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-
-%\bibliographystyle{agufull08}
-%\bibliography{sample-bib-08}
-\begin{thebibliography}{2}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-
-\bibitem[{\textit{Brown}(1967)}]{bro1967}
-Brown, R. J.~E. (1967), Permafrost in {Canada}, \textit{Map 1246A}, Geol. Surv.
-  of Can., Ottawa, Ont.
-
-\bibitem[{\textit{Englemann et~al.}(1986)\textit{Englemann, Perkins, Hagan, and
-  Haller}}]{Englemann1986}
-Englemann, R.~J., R.~W. Perkins, D.~I. Hagan, and W.~A. Haller (1986), Washout
-  coefficients for selected gases and particulates, paper presented at 59th
-  Annual Meeting, Air Pollut. Control Agency, San Francisco, Calif., 20--24
-  June.
-
-\end{thebibliography}
-
-% Reference citation examples:
-
-%...as shown by \textit{Kilby} [2008].
-%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
-%...has been shown [\textit{Kilby et al.}, 2008].
-%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
-%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
-
-%...as shown by \citet{jskilby}.
-%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
-%...has been shown \citep{jskilbye}.
-%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-
-\end{article}
-
-%----------------------------------------------------------------------------------------
-%	FIGURES AND TABLES
-%----------------------------------------------------------------------------------------
-
-%% Enter Figures and Tables here:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables; all other caption information should be placed in footnotes below the table.
-%
-%----------------
-% EXAMPLE FIGURE
-%
-% \begin{figure}
-% \noindent\includegraphics[width=20pc]{samplefigure.eps}
-% \caption{Caption text here}
-% \label{figure_label}
-% \end{figure}
-%
-% ---------------
-% EXAMPLE TABLE
-%
-%\begin{table}
-%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
-%\centering
-%\begin{tabular}{l c}
-%\hline
-% Run  & Time (min)  \\
-%\hline
-%  $l1$  & 260   \\
-%  $l2$  & 300   \\
-%  $l3$  & 340   \\
-%  $h1$  & 270   \\
-%  $h2$  & 250   \\
-%  $h3$  & 380   \\
-%  $r1$  & 370   \\
-%  $r2$  & 390   \\
-%\hline
-%\end{tabular}
-%\tablenotetext{a}{Footnote text here.}
-%\end{table}
-
-% See below for how to make sideways figures or tables.
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=0.4\linewidth]{placeholder.jpg}
-\caption{Figure caption}\label{placeholder}
-\end{figure}
-
-\begin{table}
-\caption{Table caption}\label{sampletable}
-\begin{tabular}{l l l}
-\hline
-\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
-\hline
-Treatment 1 & 0.0003262 & 0.562 \\
-Treatment 2 & 0.0015681 & 0.910 \\
-Treatment 3 & 0.0009271 & 0.296 \\
-\hline
-\end{tabular}
-\end{table}
+\begin{enumerate}
+    \item first
+    \item first
+    \item first
+    \item second
+    \item second
+\end{enumerate}
 
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
-% BREAK THE EQUATIONS INTO TWO OR MORE LINES
-% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
-% using double backslashes (\\).
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with the first character following the equal sign on the first line.
-
-% Use an \hspace{} command to insert horizontal space into your equation if necessary.
-Place an appropriate unit of measure between the curly braces, e.g.
-\hspace{1in}; you may have to experiment to achieve the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SIDEWAYS FIGURE AND TABLE EXAMPLES
-%
-%% ------------------------------------------------------------------------ %%
-%
-% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{samplefigure.eps}
-% \caption{caption here}
-% \label{label_here}
-% \end{sidewaysfigure}
-%
-% \begin{sidewaystable}
-% \caption{}
-% \begin{tabular}
-% Table layout here.
-% \end{tabular}
-% \end{sidewaystable}

### 2014-04-26 article.tex
Add bibliography back in.
Add some references and \lp pdf files.
Write 1st draft of Abstract and Intro sections, start on Methods.

@@
 
+%\newcommand{\coo}{\ce{CO2}}
+
+% 2500 words max.
+% Table of information about your simulation (as appendix): 10%
+% Discussion of climate model: 15%
+% Presentation of experimental results: 30%
+% Discussion of experiment: 30%
+% Use of wider literature: 5%
+% General structure: 10%
+%
+% 1d.p. throughout.
+% When a particular result is noted, try to discuss/explain it.
+
+% Rem: \lp
+% Analysing_your_runs
+% Converting_UM_files
+% Course_work_guide
+% geogg134merit
+% geogg134pass
+% Setting_up_your_runs
+% marengoInternal
+% WG1AR5_ALL_FINAL
+% WG1AR5_AnnexI_FINAL
+% WG1AR5_Chapter02_FINAL
+% WG1AR5_Chapter06_FINAL
+% WG1AR5_Chapter08_FINAL
+% WG1AR5_Chapter09_FINAL
+% WG1AR5_Chapter14_FINAL
+% WG1AR5_SPM_FINAL
+% WG1AR5_TS_FINAL
+% SREX-All_FINAL
+% RCP_Guide
+% SREX-Ch3-Supplement_FINAL
+
+\newif\ifsubsection
+\subsectiontrue
+%\subsectionfalse
+
 \begin{document}
 
-\part*{Using HadAM3 to model the effects of climate change}
+\part*{Using HadAM3 to model global and regional effects of increased atmospheric \ce{CO2}}
 
@@
 
-\ce{CO2} climate sensitivity (\SI{}{W.m^{-2}.K^{-1}})
+From an initial starting point of XXXppm of \ce{CO2}, HadAM3 is used to model three different scenarios: control, instantaneous doubling of \ce{CO2} and a 1\% rise in \ce{CO2} for 70 years.
+From these scenarios, the Transient Climate Response (TCR) and climate sensitivity to \ce{CO2} ($CS_{\ce{CO2}}$) are calculated, finding values of $TCR = \SI{2.04}{K}$ and $CS_{\ce{CO2}} = \SI{2.8}{W.m^{-2}.K^{-1}}$.
+The temperature differences across the globe between the control scenario and the 1\% rise scenario are analysed.
+The seasonal variation of precipitation and surface temperature over South America are discussed, and compared to the regional section of AR5 WGI IPCC (TODO: cite).
+The local changes over South America of the variables precipitation, surface temperature, Net Primary Productivity (NPP), humidity and soil moisture are analysed in detail.
+The interdependence of these variables is discussed, as well as looking into the reasons for this interdependence.
 
@@
 
-According to \textcite{gregory2004new}, you can never \dots
+% Layout the groundwork for why studying CO2 is important:
+Since 1958 the level of \ce{CO2} in the atmosphere has been monitored at the Mauna Loa observatory in Hawaii (TODO: cite).
+Its current level now stands at \SI{401}{ppm} (April, 2014), having increased from a value of \SI{315}{ppm} since 1958, which was already over the maximum value of \ce{CO2} that the Earth had seen for the previous \SI{800 000}{years} (TODO: cite).
+It is a well mixed gas, so its value at on location is indicative of its value across the globe, although there are regional variations (TODO: cite).
+It is a Greenhouse Gas (GHG), meaning that it is transparent to the majority of short-wave radiation incident on the Earth from the Sun, but opaque to the outgoing long-wave radiation, which has the effect of heating up the Earth's surface (TODO: cite Neelin).
+It is ``very likely'' that it is responsible for the recent global rise in temperatures (TODO: cite: IPCC), which in turn are predicted to lead to increased intensity of tropical cyclones (TODO: cite), droughts (TODO: cite) and extreme precipitation events (TODO: cite).
+It is therefore of vital importance that its effect on the Earth's climate is studied, so as the consequences of releasing more \ce{CO2} into the atmosphere can be predicted without performing the actual experiment.
+
+\ifsubsection
+\subsection{General Circulation Models}
+\fi
+% Explain GCMs.
+To study the release of \ce{CO2} into the atmosphere, General Circulation Models (GCMs) can be employed.
+These model the various processes that make up the Earth's dynamic behaviour such as the movements of the atmosphere or the oceans.
+They typically model these processes using a grid that spans the surface of the Earth, with variables such as temperature and pressure represented at each grid cell.
+The first GCMs just modelled the atmosphere (TODO: cite), but as they have developed they have gone on to encompass more of the Earth's systems, such as the oceans, the cryosphere and the biosphere.
+Indeed, certain large scale climatic modes such as the El ni{\~n}o Southern Oscillation (ENSO) can only be adequately modelled when the coupling between the atmosphere and the ocean is included (TODO: cite Neelin + others?).
+
+As well as modelling more of the Earth's systems, GCMs have also increased in resolution allowing smaller scale phenomena to be modelled.
+However, there are processes that occur on a scale smaller than the grid size of the models, and to include these processes they must be parameterised.
+That is, although the process is not modelled explicitly, it is included in the model by use of a parameter which describes its overall behaviour on the scale of the grid cell.
+Examples of this are cloud microphysics (TODO: check/cite) and atmospheric turbulence (TODO: check/cite).
+
+\ifsubsection
+\subsection{Radiative forcing and feedbacks}
+\fi
+% Talk about Radiative Forcing and feedbacks.
+% As \ce{CO2} increases, the incoming short-wave radiation reaching the Earth's surface is largely unaffected.
+However, more of the outgoing long-wave radiation emitted by the surface is absorbed by the atmosphere, which leads to the atmosphere heating up.
+As it heats, it too emits more long-wave radiation, but it does so isotropically, so an equal amount of this radiation is directed back towards the Earth's surface as well as into space.
+Any imbalance between the net incoming short-wave solar radiation and the net outgoing long-wave radiation from the Earth is know as a Radiative Forcing (RF), which is measure in \SI{}{W.m^{-2}}.
+A positive RF will tend to increase the temperature of the planet, whereas a negative one will decrease it.
+As the amount of \ce{CO2} in the increases, it absorbs more of the outgoing long-wave radiation, which leads to a RF.
+Similar RFs can be defined for other GHGs such as Methane.
+
+A climate feedback is a response to e.g.
+an increase in temperature that can either act to further increase the temperature (positive feedback) or decrease the temperature (negative feedback).
+One example of a climate feedback it snow albedo, which acts as a negative feedback.
+The mechanism for this is simple, as temperatures decrease, more snow can form, which increases the albedo of the surface, leading to more reflected radiation and a further drop in temperature (vice-versa for an increase).
+The most significant of these feedbacks it the water vapour feedback, which is a net positive feedback (TODO: cite).
+
+% Talk about regional effects and why they're important.
+\ifsubsection
+\subsection{Regional effects of increased \ce{CO2}}
+\fi
+% experienced at a regional level.
+A \SI{2}{K} rise in global surface temperatures cannot be directly experienced by humans; it is on a regional scale that humans will be affected.
+It is therefore important that the regional effects of increased \ce{CO2} are understood, so as plans can be put in place to mitigate the worst of these effects and changes can be made ahead of time to adapt to the changing environment.
+The predicted regional effects over South America are detailed in AR5 (TODO: cite).
+Two findings from that report will be examined using the HadAM3 model:
+\begin{itemize}
+    \item ``it is \textit{very likely} temperatures will increase over the whole continent, with greatest warming projected in southern Amazonia.''
+    \item ``It is very likely that less rainfall will occur in eastern Amazonia, northeast and eastern Brazil during the dry season.
+However, in the rainy season there is medium confidence in the precipitation changes over these regions.''
+\end{itemize}
+
+The variables within a GCM are interdependent: the precipitation in one year can affect the NPP the next year, which can affect the amount of transpiration, leading to changes in atmospheric water vapour and changes in precipitation.
+The role of these dependencies will be analysed in the context of South America, where a details consideration of five of these variables is possible as it is large enough to show spatial variation, but small enough (and geographically disconnected) to be considered as an entity.
 
@@
 \subsection{Model}
+
+HadAM3 (Hadley Centre Atmospheric Model 3) is the atmospheric component of the model HadCM3, which was developed in the late 1990s \parencite{pope2000impact}.
+It uses an Eulerian advection scheme.
+TODO:
+
+The land surface scheme is provided by the Met Office Surface Exchange Scheme (MOSES) \parencite{cox1999impact}.
+
+\subsection{Scenarios}
+
+Three scenarios are considered in this study: control, instantaneous doubling of \ce{CO2} (2x\ce{CO2}) and 1\% rise in \ce{CO2} per year (1\%py).
+The \ce{CO2} concentration for the control scenario is XXXppm, and YYYppm for the 2x\ce{CO2} and 1\%py scenarios.
+A concentration of XXXppm corresponds to 199X from the Keeling Curve.
+HadAM3 does not directly model the ocean, instead the Sea Surface Temperature (SST) was prescribed for all scenarios.
+The control and 2x\ce{CO2} scenarios were both prescribed with a SST temperature from a climate of 199X.
+For the 1\%py scenario, HadCM3 was run for \SI{80}{years} with \ce{CO2} levels increasing at 1\% per year, and the SST from this model from years 60-80 was taken as the SST for the 1\%py scenario.
+It should be noted that a 1\% increase year on year will lead to a doubling after \SI{69.7}{years}.
 
@@
 
+\printbibliography
 \appendix

### 2014-04-27 article.tex
Add some info to IPCC links.
Add in the South America correlation plots.
All figures being added to results.

@@
 % WG1AR5_AnnexI_FINAL
-% WG1AR5_Chapter02_FINAL
-% WG1AR5_Chapter06_FINAL
-% WG1AR5_Chapter08_FINAL
-% WG1AR5_Chapter09_FINAL
-% WG1AR5_Chapter14_FINAL
+% WG1AR5_Chapter02_FINAL: Observations: Atmosphere and Surface
+% WG1AR5_Chapter06_FINAL: Carbon and Other Biogeochemical Cycles
+% WG1AR5_Chapter08_FINAL: Antrhopogenic and Natural Radiative Forcing
+% WG1AR5_Chapter09_FINAL: Evaluation of Climate Models
+% WG1AR5_Chapter14_FINAL: Climate Phenomena and their Relevance for Future Regional Climate Change
 % WG1AR5_SPM_FINAL
@@
 % SREX-All_FINAL
+% ar4_wg1_full_report
 % RCP_Guide
@@
 It is a well mixed gas, so its value at on location is indicative of its value across the globe, although there are regional variations (TODO: cite).
-It is a Greenhouse Gas (GHG), meaning that it is transparent to the majority of short-wave radiation incident on the Earth from the Sun, but opaque to the outgoing long-wave radiation, which has the effect of heating up the Earth's surface (TODO: cite Neelin).
+It is a Greenhouse Gas (GHG), meaning that it is mostly transparent to the short-wave radiation incident on the Earth from the Sun, but opaque to the outgoing long-wave radiation, which has the effect of heating up the Earth's surface (TODO: cite Neelin).
 It is ``very likely'' that it is responsible for the recent global rise in temperatures (TODO: cite: IPCC), which in turn are predicted to lead to increased intensity of tropical cyclones (TODO: cite), droughts (TODO: cite) and extreme precipitation events (TODO: cite).
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/figure_1}
+    \includegraphics[width=\textwidth]{figures/global_temp_precip}
     \caption{Caption.}
-    \label{fig:figure_1}
+    \label{fig:global_temp_precip}
+\end{figure}
+
+\begin{figure}[hbp]
+    \centering
+    \includegraphics[width=\textwidth]{figures/sa_seasonal}
+    \caption{Caption.}
+    \label{fig:sa_seasonal}
+\end{figure}
+
+\begin{figure}[hbp]
+    \centering
+    \includegraphics[width=\textwidth]{figures/sa_diff}
+    \caption{Caption.}
+    \label{fig:sa_diff}
+\end{figure}
+
+\begin{figure}[hbp]
+    \centering
+    \includegraphics[width=\textwidth]{figures/corr}
+    \caption{Caption.}
+    \label{fig:corr}
 \end{figure}

### 2014-04-27 article.tex
Structure/layout of results section.

@@
 
-\newif\ifsubsection
-\subsectiontrue
-%\subsectionfalse
-
 \begin{document}
@@
 
-\ifsubsection
 \subsection{General Circulation Models}
-\fi
 % Explain GCMs.
@@
 
-\ifsubsection
 \subsection{Radiative forcing and feedbacks}
-\fi
 % Talk about Radiative Forcing and feedbacks.
@@
 % Talk about regional effects and why they're important.
-\ifsubsection
 \subsection{Regional effects of increased \ce{CO2}}
-\fi
 % experienced at a regional level.
@@
 
+\newpage
 \section{Results}
+
+\subsection{Calculation of TCR and $CS_{\ce{CO2}}$}
+
+\subsection{Global changes between 1\%py and control scenarios}
+
+Discussion about Figure \ref{fig:global_temp_precip} to go here.
 
@@
 
+\newpage
+\subsection{Seasonal changes in South America}
+
+Discussion about Figure \ref{fig:sa_seasonal} to go here.
+
 \begin{figure}[hbp]
@@
 \end{figure}
+
+\newpage
+\subsection{Interdependence of variables in South America}
+
+Discussion about Figure \ref{fig:sa_diff} to go here.
 
@@
 
+\newpage
+
+Paragraph about Figure \ref{fig:corr} plots.
 \begin{figure}[hbp]
@@
 
+\newpage
 \section{Discussion}

### 2014-04-27 article.tex
Add captions to figures, crop figures.

@@
     \includegraphics[width=\textwidth]{figures/global_temp_precip}
-    \caption{Caption.}
+    \caption{Global annual, December-February (DJF) and June-August (JJA) maps of the difference between the 1\%py and control scenarios for surface temperature and precipitation (top three left and right).
+Seasonal and zonal mean differences between the 1\% and control scenarios for temperature and precipitation (bottom two left and right).}
     \label{fig:global_temp_precip}
@@
     \includegraphics[width=\textwidth]{figures/sa_seasonal}
-    \caption{Caption.}
+    \caption{Seasonal variation of the difference between the 1\%py and control scenarios for precipitation and surface temperature for South America.}
     \label{fig:sa_seasonal}
@@
     \includegraphics[width=\textwidth]{figures/sa_diff}
-    \caption{Caption.}
+    \caption{Annual mean of the difference between the 1\%py and control scenarios for precipitation, surface temperature, humidity, NPP, and soil moisture for South America.}
     \label{fig:sa_diff}
@@
     \includegraphics[width=\textwidth]{figures/corr}
-    \caption{Caption.}
+    \caption{Scatter plots of the difference between the 1\%py and control scenarios for precipitation, surface temperature, humidity, NPP, and soil moisture for South America.
+Each pair of variables is represented once.
+Pearson's r value is shown.
+For all graphs apart from humidity-surface temperature $p < 0.001$, for humidity-surface temperature $p = 0.034$.
+}
     \label{fig:corr}

### 2014-04-27 article.tex
Note to self...
Fill in citations, start writing Results section.

--- 2014-04-27 article.tex
@@
 \usepackage{mhchem}
+\usepackage{textcomp}
 
@@
 % WG1AR5_Chapter06_FINAL: Carbon and Other Biogeochemical Cycles
-% WG1AR5_Chapter08_FINAL: Antrhopogenic and Natural Radiative Forcing
+% WG1AR5_Chapter08_FINAL: Anthropogenic and Natural Radiative Forcing
 % WG1AR5_Chapter09_FINAL: Evaluation of Climate Models
@@
 
-From an initial starting point of XXXppm of \ce{CO2}, HadAM3 is used to model three different scenarios: control, instantaneous doubling of \ce{CO2} and a 1\% rise in \ce{CO2} for 70 years.
-From these scenarios, the Transient Climate Response (TCR) and climate sensitivity to \ce{CO2} ($CS_{\ce{CO2}}$) are calculated, finding values of $TCR = \SI{2.04}{K}$ and $CS_{\ce{CO2}} = \SI{2.8}{W.m^{-2}.K^{-1}}$.
+From an initial starting point of \SI{346}{ppm} of \ce{CO2}, HadAM3 is used to model three different scenarios: control, instantaneous doubling of \ce{CO2} and a 1\% rise in \ce{CO2} for 70 years.
+From these scenarios, the Transient Climate Response (TCR) and climate sensitivity to \ce{CO2} ($CS_{\ce{CO2}}$) are calculated, finding values of $TCR = \SI{2.0}{K}$ and $CS_{\ce{CO2}} = \SI{2.9}{W.m^{-2}.K^{-1}}$.
 The temperature differences across the globe between the control scenario and the 1\% rise scenario are analysed.
-The seasonal variation of precipitation and surface temperature over South America are discussed, and compared to the regional section of AR5 WGI IPCC (TODO: cite).
+The seasonal variation of precipitation and surface temperature over South America are discussed, and compared to the regional section of IPCC AR5 WG I \parencite{ipcc2014wg1}.
 The local changes over South America of the variables precipitation, surface temperature, Net Primary Productivity (NPP), humidity and soil moisture are analysed in detail.
@@
 % Layout the groundwork for why studying CO2 is important:
-Since 1958 the level of \ce{CO2} in the atmosphere has been monitored at the Mauna Loa observatory in Hawaii (TODO: cite).
-Its current level now stands at \SI{401}{ppm} (April, 2014), having increased from a value of \SI{315}{ppm} since 1958, which was already over the maximum value of \ce{CO2} that the Earth had seen for the previous \SI{800 000}{years} (TODO: cite).
-It is a well mixed gas, so its value at on location is indicative of its value across the globe, although there are regional variations (TODO: cite).
-It is a Greenhouse Gas (GHG), meaning that it is mostly transparent to the short-wave radiation incident on the Earth from the Sun, but opaque to the outgoing long-wave radiation, which has the effect of heating up the Earth's surface (TODO: cite Neelin).
-It is ``very likely'' that it is responsible for the recent global rise in temperatures (TODO: cite: IPCC), which in turn are predicted to lead to increased intensity of tropical cyclones (TODO: cite), droughts (TODO: cite) and extreme precipitation events (TODO: cite).
+Since 1958 the level of \ce{CO2} in the atmosphere has been monitored at the Mauna Loa observatory in Hawaii \parencite{keeling1976atmospheric}.
+Its current level now stands at \SI{401}{ppm} (April, 2014), having increased from a value of \SI{315}{ppm} since 1958, which was already over the maximum value of \ce{CO2} that the Earth had seen for the previous \SI{800 000}{years} \parencite{yin2012individual}.
+It is a well mixed gas, so its value at on location is indicative of its value across the globe, although there are regional variations \parencite{gurney2002towards}.
+It is a Greenhouse Gas (GHG), meaning that it is mostly transparent to the short-wave radiation incident on the Earth from the Sun, but opaque to the outgoing long-wave radiation, which has the effect of heating up the Earth's surface \parencite{neelin2011climate}.
+It is ``very likely'' that it is responsible for the recent global rise in temperatures \parencite{ipcc2014wg1}, which in turn are predicted to lead to increased intensity of tropical cyclones \parencite{knutson2010tropical}, droughts \parencite{dai2012increasing} and extreme precipitation events \parencite{shongwe2009projected}.
 It is therefore of vital importance that its effect on the Earth's climate is studied, so as the consequences of releasing more \ce{CO2} into the atmosphere can be predicted without performing the actual experiment.
 
-\ifsubsection
 \subsection{General Circulation Models}
-\fi
 % Explain GCMs.
@@
 They typically model these processes using a grid that spans the surface of the Earth, with variables such as temperature and pressure represented at each grid cell.
-The first GCMs just modelled the atmosphere (TODO: cite), but as they have developed they have gone on to encompass more of the Earth's systems, such as the oceans, the cryosphere and the biosphere.
-Indeed, certain large scale climatic modes such as the El ni{\~n}o Southern Oscillation (ENSO) can only be adequately modelled when the coupling between the atmosphere and the ocean is included (TODO: cite Neelin + others?).
+Early GCMs just modelled the atmosphere \parencite{holloway1971simulation}, but as they have developed they have gone on to encompass more of the Earth's systems, such as the oceans, the cryosphere and the biosphere.
+Indeed, certain large scale climatic modes such as the El ni{\~n}o Southern Oscillation (ENSO) can only be adequately modelled when the coupling between the atmosphere and the ocean is included \parencite{neelin2011climate}.
 
@@
 That is, although the process is not modelled explicitly, it is included in the model by use of a parameter which describes its overall behaviour on the scale of the grid cell.
-Examples of this are cloud microphysics (TODO: check/cite) and atmospheric turbulence (TODO: check/cite).
+An example of this are the parameterisation of cloud microphysics \parencite{walko1995new, neelin2011climate}.
 
-\ifsubsection
 \subsection{Radiative forcing and feedbacks}
-\fi
 % Talk about Radiative Forcing and feedbacks.
@@
 
-Three scenarios are considered in this study: control, instantaneous doubling of \ce{CO2} (2x\ce{CO2}) and 1\% rise in \ce{CO2} per year (1\%py).
-The \ce{CO2} concentration for the control scenario is XXXppm, and YYYppm for the 2x\ce{CO2} and 1\%py scenarios.
+Three scenarios are considered in this study: control, instantaneous doubling of \ce{CO2} (2x\ce{CO2}) and 1\% rise in \ce{CO2} per year (1\%/yr).
+The \ce{CO2} concentration for the control scenario is XXXppm, and YYYppm for the 2x\ce{CO2} and 1\%/yr scenarios.
 A concentration of XXXppm corresponds to 199X from the Keeling Curve.
@@
 The control and 2x\ce{CO2} scenarios were both prescribed with a SST temperature from a climate of 199X.
-For the 1\%py scenario, HadCM3 was run for \SI{80}{years} with \ce{CO2} levels increasing at 1\% per year, and the SST from this model from years 60-80 was taken as the SST for the 1\%py scenario.
+For the 1\%/yr scenario, HadCM3 was run for \SI{80}{years} with \ce{CO2} levels increasing at 1\% per year, and the SST from this model from years 60-80 was taken as the SST for the 1\%/yr scenario.
 It should be noted that a 1\% increase year on year will lead to a doubling after \SI{69.7}{years}.
 
+\newpage
 \section{Results}
+
+\subsection{Calculation of TCR and $CS_{\ce{CO2}}$}
+
+\subsection{Global changes between 1\%/yr and control scenarios}
+
+% Points to mention:
+% Temp: Highest over Arctic + strongest effect seen in winter. Warning over Antarctic in its winter is far less pronounced.
+% only some areas cool at all
+
+% Precip: very different spatial pattern to temp, interesting effect either side of eq. , plus large red. over amazonia DJF visible. Strongest effect in Nov.. 
+Figure \ref{fig:global_temp_precip} top-left shows the difference between the 1\%/yr and control scenarios for surface temperature averaged over the year.
+From this map it is clear that the 1\%/yr scenario is the same temperature or hotter across almost the whole globe.
+There is only a small area in the Southern Ocean (centred on 60\textdegree S, 120\textdegree E) where the 1\%/yr scenario is cooler.
+This reflects the value of TCR of \SI{2}{K} found in the previous section.
+It is also clear from this map that the Arctic shows by far the biggest difference, and this difference is only increased when looking at the northern hemisphere winter December-February (DJF) map, where the difference between the scenarios can reach \SI{20}{K}.
+However the similar southern hemisphere winter map shows that warming over the Antarctic is far less pronounced.
+The seasonal dependence of the warming magnitude can be seen in the ``Seasonal mean temperature'' graph, where warming can be seen to be greatest from September to January.
+The ``Zonal mean temperature'' graph confirms that the heating happens predominantly over the Arctic, with warming being far larger at latitudes higher than 60\textdegree N.
+
+The difference in precipitation between the two scenarios shows a different spatial pattern.
+In the ``Annual'' map, the main increases can be seen to happen over the Pacific Atlantic and Indian Oceans, just north of the Equator.
+In the Indian Ocean, there is another area of large increase that can be seen just south of the Equator, and this can be seen as a prominent feature in the ``DJF'' map.
+The main areas of precipitation reduction are the eastern Pacific Ocean, and over northern Brazil.
+The reduction over northern Brazil is most prominent in the ``DJF'' map, suggesting a seasonal dependence in this reduction.
+The ``Seasonal mean precip.'' graph shows a strong increase in November.
+The ``Zonal mean precip.'' graph shows an increase just north of the Equator, and a corresponding decrease just south of the Equator.
+There is a broad increase around 50°S as well.
+% TODO: Explain!
 
@@
     \includegraphics[width=\textwidth]{figures/global_temp_precip}
-    \caption{Caption.}
+    \caption{Global annual, December-February (DJF) and June-August (JJA) maps of the difference between the 1\%/yr and control scenarios for surface temperature and precipitation (top three left and right).
+Seasonal and zonal mean differences between the 1\% and control scenarios for temperature and precipitation (bottom two left and right).}
     \label{fig:global_temp_precip}
 \end{figure}
+
+\newpage
+\subsection{Seasonal changes in South America}
+
+% Points to mention:
+% precip: Large reduction over amazon in DJF. How this compares to IPCC.
+% Large increase over Amazon/North coast of Brazil across the year. Signs of correlation.
+From Figure \ref{fig:sa_seasonal}, DJF $\Delta$ Precip.
+(top left), the same reduction in precipitation over the mouth of the Amazon in northern Brazil that was mentioned in the previous section is clearly visible.
+This reduction persists across the course of the year (as confirmed in Figure \ref{fig:sa_diff}, top left), although it is most pronounced in DJF.
+When looking at the change in surface temperature over the same region, the whole of the continent is seen so undergo substantial heating.
+The average temperature increase over the land is \SI{3.4}{K}.
+The heating is, again, most prominent over northern Brazil, at the mouth of the Amazon.
 
@@
     \includegraphics[width=\textwidth]{figures/sa_seasonal}
-    \caption{Caption.}
+    \caption{Seasonal variation of the difference between the 1\%/yr and control scenarios for precipitation and surface temperature for South America.}
     \label{fig:sa_seasonal}
 \end{figure}
+
+\newpage
+
+\subsection{Interdependence of variables in South America}
+
+% Points to mention:
+% Area most affected for 4/5 seems to be north coast of Br., signs of correlation between these 5 variables. 
+In the five variables shown in Figure \ref{fig:sa_diff}, three show a broad continental increase between the 1\%/yr and control scenarios: surface temperature, humidity and NPP.
+Conversely, precipitation and soil moisture both show decreases over the continent.
+What is interesting is that each of the variables shows an increase or decrease over the head of the Amazon, which hints at these variables' interdependence on each other.
+That some these variables should be linked is not surprising: it makes sense \textit{a priori} that if e.g.
+precipitation goes up between the two scenarios, soil moisture should go up as well.
 
@@
     \includegraphics[width=\textwidth]{figures/sa_diff}
-    \caption{Caption.}
+    \caption{Annual mean of the difference between the 1\%/yr and control scenarios for precipitation, surface temperature, humidity, NPP, and soil moisture for South America.}
     \label{fig:sa_diff}
 \end{figure}
+
+\newpage
+
+% Points to mention:
+% Most plots show some amount of correlation (apart from humidity/surf temp.). Implies an interdependency between these variables. explain plausible mechanism for some of these interdependencies. 
+This interdependence is examined further in Figure \ref{fig:corr}, which plots the difference of each of the variables between the 1\%/yr and control scenarios against every other variable's difference, showing Pearson's correlation statistic for each pair.
+This shows that when e.g.
+there is a strong correlation between the increase in precipitation and the increase in soil moisture between the two scenarios.
 
@@
     \includegraphics[width=\textwidth]{figures/corr}
-    \caption{Caption.}
+    \caption{Scatter plots of the difference between the 1\%/yr and control scenarios for precipitation, surface temperature, humidity, NPP, and soil moisture for South America.
+Each pair of variables is represented once.
+Pearson's r value is shown.
+For all graphs apart from humidity-surface temperature $p < 0.001$, for humidity-surface temperature $p = 0.034$.
+}
     \label{fig:corr}

### 2014-04-27 article.tex
Fill in most of appendix with required info.
Finish of appendix.

@@
 
-\section{Required output}
+\section{Additional information}
+
+N.B. more decimal points quoted than required for detailed numerical comparison.
 
 \begin{enumerate}
-    \item first
-    \item first
-    \item first
-    \item second
-    \item second
+    \item \textbf{username on \texttt{puma.nerc.ac.uk}}: UCL3
+    \item \textbf{simulation id}: xjksa
+    \item \textbf{simulation scenario}: 2x\ce{CO2}
+    \item \textbf{\ce{CO2} level of scenario}: \SI{692.1}{ppm}
+    \item \textbf{global annual average surface temperature}: \SI{277.859}{K}
+    \item \textbf{$G_{2x\ce{CO2}}$}: \SI{3.71}{W.m^{-2}}
+    \item \textbf{global annual average surface net flux in control}: \SI{10.102}{W.m^{-2}}
+    \item \textbf{Transient Climate Response (TCR)}: \SI{2.044}{K}
+    \item \textbf{$\Delta T^{eff}_{2x\ce{CO2}}$}: \SI{2.85}{W.m^{-2}.K^{-1}}
+    \item \textbf{RMS surface temperature bias in DJF from control}: \SI{3.943}{K} % TODO: not sure about this one at all.
 \end{enumerate}

### 2014-04-27 article.tex
Fill in results section on TCR/clim. sens.

@@
 From an initial starting point of \SI{346}{ppm} of \ce{CO2}, HadAM3 is used to model three different scenarios: control, instantaneous doubling of \ce{CO2} and a 1\% rise in \ce{CO2} for 70 years.
-From these scenarios, the Transient Climate Response (TCR) and climate sensitivity to \ce{CO2} ($CS_{\ce{CO2}}$) are calculated, finding values of $TCR = \SI{2.0}{K}$ and $CS_{\ce{CO2}} = \SI{2.9}{W.m^{-2}.K^{-1}}$.
+From these scenarios, the Transient Climate Response (TCR) and effective climate sensitivity to \ce{CO2} ($\Delta T^{eff}_{\ce{CO2}}$, ) are calculated, finding values of $TCR = \SI{2.0}{K}$ and $\Delta T^{eff}_{\ce{CO2}} = \SI{2.9}{K}$.
 The temperature differences across the globe between the control scenario and the 1\% rise scenario are analysed.
@@
 It is therefore important that the regional effects of increased \ce{CO2} are understood, so as plans can be put in place to mitigate the worst of these effects and changes can be made ahead of time to adapt to the changing environment.
-The predicted regional effects over South America are detailed in AR5 (TODO: cite).
+The predicted regional effects over South America are detailed in IPCC AR5 WG I \parencite{ipcc2014wg1}.
 Two findings from that report will be examined using the HadAM3 model:
@@
 The variables within a GCM are interdependent: the precipitation in one year can affect the NPP the next year, which can affect the amount of transpiration, leading to changes in atmospheric water vapour and changes in precipitation.
-The role of these dependencies will be analysed in the context of South America, where a details consideration of five of these variables is possible as it is large enough to show spatial variation, but small enough (and geographically disconnected) to be considered as an entity.
+The role of these dependencies will be analysed in the context of South America, where a details consideration of five of these variables is possible as it is large enough to show spatial variation, but small enough (and geographically disconnected) to be considered as a discrete entity.
 
@@
 Three scenarios are considered in this study: control, instantaneous doubling of \ce{CO2} (2x\ce{CO2}) and 1\% rise in \ce{CO2} per year (1\%/yr).
-The \ce{CO2} concentration for the control scenario is XXXppm, and YYYppm for the 2x\ce{CO2} and 1\%/yr scenarios.
-A concentration of XXXppm corresponds to 199X from the Keeling Curve.
+The \ce{CO2} concentration for the control scenario is \SI{374}, and \SI{748} for the 2x\ce{CO2} and 1\%/yr scenarios.
+A concentration of \SI{374} corresponds to approximately 2005 from the Keeling Curve \parencite{keeling1976atmospheric}.
 HadAM3 does not directly model the ocean, instead the Sea Surface Temperature (SST) was prescribed for all scenarios.
@@
 
-\subsection{Calculation of TCR and $CS_{\ce{CO2}}$}
+\subsection{Calculation of TCR and $\Delta T_{\ce{CO2}}$}
+
+Once three scenarios (control, 1\%/yr and 2x\ce{CO2}) have been run, some global statistics can be calculated that are useful when comparing the results of different models.
+The first of these is the TCR, which is the difference in global mean temperature between the 1\%/yr and the control scenarios.
+The calculated value for HadAM3 in this study is TCR = \SI{2.0}{K}.
+This is in agreement with IPCC AR5 WG I \parencite{ipcc2014wg1}, which finds that the range is 1.0 to \SI{2.5}{K}.
+
+Another useful parameter in comparing the size of response for different models is the climate sensitivity due to \ce{CO2} ($\Delta T_{\ce{CO2}}$), or the surface temperature change that you would see if you doubled \ce{CO2} concentration and let the Earth reach a new equilibrium.
+Following method B of \textcite{gregory2004new}, the ``effective climate sensitivity'', $\Delta T^{eff}_{\ce{CO2}}$, can be calculated from the TCR and two flux imbalances, one from each of the 2x\ce{CO2} and 1\%/yr scenarios.
+If the imposed forcing of doubling \ce{CO2} is $F$, and the net forcing after a given time is $N$, then the climate feedback parameter $\alpha$ can be calculated using:
+
+\begin{equation}
+    \alpha = \frac{(F - N)}{\Delta T}
+    \label{eqn:alpha}
+\end{equation}
+
+Here $\Delta T$ is the TCR already calculated, $F$ is calculated from the 2x\ce{CO2} scenario by taking the difference of the net top of atmosphere flux between the scenario and the control scenario.
+Similarly, $N$ is calculated from the 1\%/yr scenario, which leads to values of $F = \SI{3.71}{W.m^{-2}}$, and $N = \SI{1.05}{W.m^{-2}}$.
+This gives $\alpha = \SI{1.30}{W.m^{-2}.K^{-1}}$.
+Finally, the equation for climate sensitivity is:
+
+\begin{equation}
+    \Delta T^{eff}_{\ce{CO2}} = \frac{F}{\alpha}
+    \label{eqn:delta_t}
+\end{equation}
+
+Using this equation with the above number yields $\Delta T^{eff}_{\ce{CO2}} = \SI{2.85}{K}$.
+This is in reasonable agreement with the range found in IPCC AR5 WG I, which gives the range as 1.5 to \SI{4.5}{K} \parencite{ipcc2014wg1}.
 
@@
 \newpage
+
 \section{Discussion}
@@
     \item \textbf{Transient Climate Response (TCR)}: \SI{2.044}{K}
-    \item \textbf{$\Delta T^{eff}_{2x\ce{CO2}}$}: \SI{2.85}{W.m^{-2}.K^{-1}}
+    \item \textbf{$\Delta T^{eff}_{2x\ce{CO2}}$}: \SI{2.85}{K}
     \item \textbf{RMS surface temperature bias in DJF from control}: \SI{3.943}{K} % TODO: not sure about this one at all.

### 2014-04-27 article.tex
Re-arrange some of the Intro to later, write concl.
Take out an unneeded secion on feedbacks.

@@
 From these scenarios, the Transient Climate Response (TCR) and effective climate sensitivity to \ce{CO2} ($\Delta T^{eff}_{\ce{CO2}}$, ) are calculated, finding values of $TCR = \SI{2.0}{K}$ and $\Delta T^{eff}_{\ce{CO2}} = \SI{2.9}{K}$.
-The temperature differences across the globe between the control scenario and the 1\% rise scenario are analysed.
-The seasonal variation of precipitation and surface temperature over South America are discussed, and compared to the regional section of IPCC AR5 WG I \parencite{ipcc2014wg1}.
+The surface temperature and precipitation differences across the globe between the control scenario and the 1\% rise scenario are analysed.
+The seasonal variation of precipitation and surface temperature over South America are analysed, and compared to the regional section of IPCC AR5 WG I \parencite{ipcc2014wg1}.
 The local changes over South America of the variables precipitation, surface temperature, Net Primary Productivity (NPP), humidity and soil moisture are analysed in detail.
-The interdependence of these variables is discussed, as well as looking into the reasons for this interdependence.
+The interdependence of these variables is demonstrated.
+% TODO, as well as looking into the reasons for this interdependence.
 
@@
 
-\subsection{General Circulation Models}
+\subsection{General circulation models}
 % Explain GCMs.
@@
 
-\subsection{Radiative forcing and feedbacks}
-% Talk about Radiative Forcing and feedbacks.
-% As \ce{CO2} increases, the incoming short-wave radiation reaching the Earth's surface is largely unaffected.
-However, more of the outgoing long-wave radiation emitted by the surface is absorbed by the atmosphere, which leads to the atmosphere heating up.
-As it heats, it too emits more long-wave radiation, but it does so isotropically, so an equal amount of this radiation is directed back towards the Earth's surface as well as into space.
+\subsection{Radiative forcing}
+% Talk about Radiative Forcing / no longer! and feedbacks.
 Any imbalance between the net incoming short-wave solar radiation and the net outgoing long-wave radiation from the Earth is know as a Radiative Forcing (RF), which is measure in \SI{}{W.m^{-2}}.
 A positive RF will tend to increase the temperature of the planet, whereas a negative one will decrease it.
-As the amount of \ce{CO2} in the increases, it absorbs more of the outgoing long-wave radiation, which leads to a RF.
+As the amount of \ce{CO2} in the atmosphere increases, it absorbs more of the outgoing long-wave radiation, which leads to a RF.
 Similar RFs can be defined for other GHGs such as Methane.
 
-A climate feedback is a response to e.g.
+%A climate feedback is a response to e.g.
 an increase in temperature that can either act to further increase the temperature (positive feedback) or decrease the temperature (negative feedback).
-One example of a climate feedback it snow albedo, which acts as a negative feedback.
-The mechanism for this is simple, as temperatures decrease, more snow can form, which increases the albedo of the surface, leading to more reflected radiation and a further drop in temperature (vice-versa for an increase).
-The most significant of these feedbacks it the water vapour feedback, which is a net positive feedback (TODO: cite).
+One example of a climate feedback is snow albedo, which acts as a negative feedback.
+The mechanism for this is simple: as temperatures decrease, more snow can form, which increases the albedo of the surface, leading to more reflected radiation and a further drop in temperature (vice-versa for an increase).
+The most significant of these feedbacks it the water vapour feedback, which is a net positive feedback.
 
@@
 \subsection{Regional effects of increased \ce{CO2}}
-% experienced at a regional level.
+\label{sec:regional_effects}
 A \SI{2}{K} rise in global surface temperatures cannot be directly experienced by humans; it is on a regional scale that humans will be affected.
-It is therefore important that the regional effects of increased \ce{CO2} are understood, so as plans can be put in place to mitigate the worst of these effects and changes can be made ahead of time to adapt to the changing environment.
+It is therefore important that the regional effects of increased \ce{CO2} are understood, so that plans can be put in place to mitigate the worst of these effects and changes can be made ahead of time to adapt to the changing environment.
 The predicted regional effects over South America are detailed in IPCC AR5 WG I \parencite{ipcc2014wg1}.
-Two findings from that report will be examined using the HadAM3 model:
-\begin{itemize}
-    \item ``it is \textit{very likely} temperatures will increase over the whole continent, with greatest warming projected in southern Amazonia.''
-    \item ``It is very likely that less rainfall will occur in eastern Amazonia, northeast and eastern Brazil during the dry season.
-However, in the rainy season there is medium confidence in the precipitation changes over these regions.''
-\end{itemize}
+Two findings from that report will be examined in Section \ref{subsec:seasonal_changes}.
 
-The variables within a GCM are interdependent: the precipitation in one year can affect the NPP the next year, which can affect the amount of transpiration, leading to changes in atmospheric water vapour and changes in precipitation.
+The variables within a GCM are interdependent: the amount of precipitation can affect the NPP, which in turn can affect the amount of transpiration, leading to changes in atmospheric water vapour and changes in precipitation.
 The role of these dependencies will be analysed in the context of South America, where a details consideration of five of these variables is possible as it is large enough to show spatial variation, but small enough (and geographically disconnected) to be considered as a discrete entity.
@@
 It uses an Eulerian advection scheme.
-TODO:
+An Arakawa B grid is used, and the horizontal resolution is 2.5\textdegree\,  latitude by 3.5\textdegree\, longitude, which corresponds to grid cells of approximately \SI{280}{km} by \SI{390}{km} at the equator.
+There are 19 vertical levels and the model employs a hydrostatic vertical coordinate.
+The model is run on a \SI{30}{minute} timestep.
 
 The land surface scheme is provided by the Met Office Surface Exchange Scheme (MOSES) \parencite{cox1999impact}.
+MOSES includes a representation of freezing and melting of the soil moisture layer, which leads to better simulation of surface temperatures.
+It also includes a new formulation of evaporation, which models the dependence of stomatal resistance on temperature, vapour pressure deficit and \ce{CO2}.
 
@@
 The \ce{CO2} concentration for the control scenario is \SI{374}, and \SI{748} for the 2x\ce{CO2} and 1\%/yr scenarios.
-A concentration of \SI{374} corresponds to approximately 2005 from the Keeling Curve \parencite{keeling1976atmospheric}.
+A concentration of \SI{374}{ppm} corresponds to approximately 2005 from the Keeling Curve \parencite{keeling1976atmospheric}.
 HadAM3 does not directly model the ocean, instead the Sea Surface Temperature (SST) was prescribed for all scenarios.
-The control and 2x\ce{CO2} scenarios were both prescribed with a SST temperature from a climate of 199X.
+The control and 2x\ce{CO2} scenarios were both prescribed with a SST corresponding to a contemporary climate.
 For the 1\%/yr scenario, HadCM3 was run for \SI{80}{years} with \ce{CO2} levels increasing at 1\% per year, and the SST from this model from years 60-80 was taken as the SST for the 1\%/yr scenario.
 It should be noted that a 1\% increase year on year will lead to a doubling after \SI{69.7}{years}.
+Each of the runs was run for \SI{21}{years} and \SI{1}{month}.
 
@@
 
-\subsection{Calculation of TCR and $\Delta T_{\ce{CO2}}$}
+\subsection{Calculation of TCR and $\Delta T^{eff}_{\ce{CO2}}$}
 
@@
 Here $\Delta T$ is the TCR already calculated, $F$ is calculated from the 2x\ce{CO2} scenario by taking the difference of the net top of atmosphere flux between the scenario and the control scenario.
-Similarly, $N$ is calculated from the 1\%/yr scenario, which leads to values of $F = \SI{3.71}{W.m^{-2}}$, and $N = \SI{1.05}{W.m^{-2}}$.
-This gives $\alpha = \SI{1.30}{W.m^{-2}.K^{-1}}$.
-Finally, the equation for climate sensitivity is:
+$N$ is calculated in the same way from the 1\%/yr scenario, which leads to values of $F = \SI{3.71}{W.m^{-2}}$, and $N = \SI{1.05}{W.m^{-2}}$.
+This gives a value of $\alpha = \SI{1.30}{W.m^{-2}.K^{-1}}$.
+The equation for climate sensitivity is:
 
@@
 The ``Zonal mean precip.'' graph shows an increase just north of the Equator, and a corresponding decrease just south of the Equator.
-There is a broad increase around 50°S as well.
+There is a broad increase around 50\textdegree S as well.
 % TODO: Explain!
@@
 \subsection{Seasonal changes in South America}
+\label{subsec:seasonal_changes}
 
@@
 This reduction persists across the course of the year (as confirmed in Figure \ref{fig:sa_diff}, top left), although it is most pronounced in DJF.
-When looking at the change in surface temperature over the same region, the whole of the continent is seen so undergo substantial heating.
+When looking at the change in surface temperature over the same region, the whole of the continent is seen to undergo substantial heating.
 The average temperature increase over the land is \SI{3.4}{K}.
@@
 \end{figure}
+
+Two findings from IPCC AR5 WG I \parencite{ipcc2014wg1} are:
+\begin{enumerate}
+    \item ``it is \textit{very likely} temperatures will increase over the whole continent, with greatest warming projected in southern Amazonia.''
+    \item ``It is very likely that less rainfall will occur in eastern Amazonia, northeast and eastern Brazil during the dry season.
+However, in the rainy season there is medium confidence in the precipitation changes over these regions.''
+\end{enumerate}
+
+From the maps in Figure \ref{fig:sa_seasonal}, it is clear that the temperature increases over the first whole continent.
+Therefore The first part of the first finding is replicated by HadAM3.
+However, the most warming is seen in the north of the Amazon, not the south as in the IPCC report.
+This could be because the IPCC report uses higher resolution models, that can better reproduce the spatial variation over continents.
+It could also be because the IPCC report uses regional models to assist in the modelling of certain regions, which will provide more detail (and it is hoped more accuracy) over specific regions of the Earth.
+
+For the second finding, the dry season in Brazil is best shown by the SON map, which does indeed show a significant reduction in precipitation over eastern Brazil.
+However, a higher reduction is seen in the DJF map, suggesting that once more the HadAM3 model may not be producing the interior detail of the South American continent.
 
@@
 precipitation goes up between the two scenarios, soil moisture should go up as well.
+The variables humidity and soil moisture have more than one vertical value, so the mean over the vertical coordinate was taken.
 
@@
 
+The HadAM3 model is
+
 \section{Conclusions}
+
+Although HadAM3 could not match the predictions of more recent models (as in Section \ref{subsec:seasonal_changes}), it has been used to produce some results that are consistent with the current values, particularly the TCR and the $\Delta T^{eff}_{2x\ce{CO2}}$.
+Given the low computational cost of running this model - running the model for \SI{21}{years} took about 5 days on a standard computer - makes it well suited for running in a distributed manner across commodity hardware,  as is  being done by climateprediction.net \parencite{stainforth2005uncertainty}.
+Also, as was done in this study, it can be used to shed light on the interdependence of various variables.
 

### 2014-04-27 article.tex
Add a discussion.

@@
 
-The HadAM3 model is
+This study has identified some correlations between variables within South America.
+However, all mention of causality has been carefully avoided, as there is not enough evidence to say that e.g.
+a decrease in precipitation has lead to a decrease in NPP in South America.
+A further line of enquiry could be to look for causal links between the variables that have been correlated together here.
+To do this, it is likely that the data would have to be analysed without taking temporal means, so as it could be established if the change in one particular variable preceeded the change in another (implying causation).
+Alternatively, the model itself could be modified so as certain aspects of the land surface scheme, MOSES, could be turned on or off, as is done currently for the radiation scheme currently \parencite{pope2000impact}.
 

### 2014-04-28 article.tex
Add coversheet, small cleanup.

@@
 \lhead{}
-\chead{BTTM8 - CLIMATE MODELLING - USING HADAM3 TO MODEL CLIMATE CHANGE}
+\chead{}
 \rhead{}
@@
 
+\textbf{UCL DEPARTMENT OF GEOGRAPHY}\\[4cm]
+
+\LARGE
+
+\textbf{EXAM CANDIDATE ID:} BTTM8
+
+\textbf{DEPARTMENT/COLLEGE:} Department of Geography
+
+\textbf{MODULE CODE:} GEOGG134
+
+\textbf{MODULE TITLE:} Climate Modelling
+
+\textbf{COURSE PAPER TITLE:} Using HadAM3 to model global and regional effects of increased atmospheric \ce{CO2} \\[3cm]
+
+\textbf{NAME OF ACADEMIC TUTOR:} Chris Brierley
+
+\textbf{WORD COUNT:} 2448
+
+\normalsize
+
+\newpage
+
+\chead{BTTM8 - CLIMATE MODELLING - USING HADAM3 TO MODEL CLIMATE CHANGE}
+
 \part*{Using HadAM3 to model global and regional effects of increased atmospheric \ce{CO2}}
@@
 % Talk about Radiative Forcing / no longer! and feedbacks.
-Any imbalance between the net incoming short-wave solar radiation and the net outgoing long-wave radiation from the Earth is know as a Radiative Forcing (RF), which is measure in \SI{}{W.m^{-2}}.
+Any imbalance between the net incoming short-wave solar radiation and the net outgoing long-wave radiation from the Earth is know as a Radiative Forcing (RF), which is measured in \SI{}{W.m^{-2}}.
 A positive RF will tend to increase the temperature of the planet, whereas a negative one will decrease it.
@@
 Three scenarios are considered in this study: control, instantaneous doubling of \ce{CO2} (2x\ce{CO2}) and 1\% rise in \ce{CO2} per year (1\%/yr).
-The \ce{CO2} concentration for the control scenario is \SI{374}, and \SI{748} for the 2x\ce{CO2} and 1\%/yr scenarios.
+The \ce{CO2} concentration for the control scenario is \SI{374}{ppm}, and \SI{748}{ppm} for the 2x\ce{CO2} and 1\%/yr scenarios.
 A concentration of \SI{374}{ppm} corresponds to approximately 2005 from the Keeling Curve \parencite{keeling1976atmospheric}.
@@
 It should be noted that a 1\% increase year on year will lead to a doubling after \SI{69.7}{years}.
-Each of the runs was run for \SI{21}{years} and \SI{1}{month}.
+Each of the scenarios was run for \SI{21}{years} and \SI{1}{month}.
 
@@
 Using this equation with the above number yields $\Delta T^{eff}_{\ce{CO2}} = \SI{2.85}{K}$.
-This is in reasonable agreement with the range found in IPCC AR5 WG I, which gives the range as 1.5 to \SI{4.5}{K} \parencite{ipcc2014wg1}.
+This is in agreement with the range found in IPCC AR5 WG I, which gives the range as 1.5 to \SI{4.5}{K} \parencite{ipcc2014wg1}.
 
@@
 This reflects the value of TCR of \SI{2}{K} found in the previous section.
-It is also clear from this map that the Arctic shows by far the biggest difference, and this difference is only increased when looking at the northern hemisphere winter December-February (DJF) map, where the difference between the scenarios can reach \SI{20}{K}.
+It is also clear from this map that the Arctic shows by far the biggest difference, and this difference is only increased when looking at the northern hemisphere winter December-February (DJF) map, where the difference between the scenarios can reach over \SI{20}{K}.
 However the similar southern hemisphere winter map shows that warming over the Antarctic is far less pronounced.
@@
 % Most plots show some amount of correlation (apart from humidity/surf temp.). Implies an interdependency between these variables. explain plausible mechanism for some of these interdependencies. 
-This interdependence is examined further in Figure \ref{fig:corr}, which plots the difference of each of the variables between the 1\%/yr and control scenarios against every other variable's difference, showing Pearson's correlation statistic for each pair.
+This interdependence is examined further in Figure \ref{fig:corr}, which plots the difference of each of the variables between the 1\%/yr and control scenarios against every other variable's difference, showing Pearson's correlation coefficient for each pair.
 This shows that when e.g.
@@
 To do this, it is likely that the data would have to be analysed without taking temporal means, so as it could be established if the change in one particular variable preceeded the change in another (implying causation).
-Alternatively, the model itself could be modified so as certain aspects of the land surface scheme, MOSES, could be turned on or off, as is done currently for the radiation scheme currently \parencite{pope2000impact}.
+Alternatively, the model itself could be modified so as certain aspects of the land surface scheme, MOSES, could be turned on or off, as is currently done for the radiation scheme \parencite{pope2000impact}.
 
